%\begingroup
%\let\cleardoublepage\clearpage

% This ensures that only the main file is compiled
% !TEX root = ../my_thesis.tex

% English abstract
\chapter*{List of publications}
\markboth{List of publications}{List of publications}
%\addcontentsline{toc}{chapter}{List of publications} % adds an entry to the table of contents
\vskip0.5cm
% put your text here

%First part of Chapter \ref{ch:es_literature} is built on an opinion published in a journal:

\begin{quote}
Limpens, G., \textbf{Rixhon, X.}, Contino, F., \& Jeanmart, H. (2024). ``\emph{EnergyScope Pathway: An open-source model to optimise the energy transition pathways of a regional whole-energy system.}'' In Applied Energy, (Vol. 358). URL: \url{https://doi.org/10.1016/j.apenergy.2023.122501}
\end{quote}

\begin{quote}
\textbf{Rixhon, X.}, Limpens, G., Coppitters, D., Jeanmart, H.,\& Contino, F.(2022). ``\emph{The role of electrofuels under uncertainties for the Belgian energy transition.}'' In Energies (Vol. 14). URL: \url{https://doi.org/10.3390/en14134027}
\end{quote}

\begin{quote}
\textbf{Rixhon, X.}, Limpens, G., Contino, F., \& Jeanmart, H. (2021). ``\emph{Taxonomy of the fuels in a whole-energy system.}'' In Frontiers in Energy Research, Sec. Sustainable Energy Systems, (Vol. 9). URL: \url{https://doi.org/10.3389/fenrg.2021.660073} 
\end{quote}

\begin{quote}
\textbf{Rixhon, X.}, Tonelli, D., Colla, M., Verleysen, K., Limpens, G., Jeanmart, H. ,\& Contino, F.(2022). ``\emph{Integration of non-energy among the end-use demands of bottom-up whole-energy system models.}'' In Frontiers in Energy Research, Sec. Process and Energy Systems Engineering, (Vol. 10). URL: \url{https://doi.org/10.3389/fenrg.2022.904777}
\end{quote}

\begin{quote}
\textbf{Rixhon, X.}, Colla, M., Tonelli, D., Verleysen, K.,  Limpens, G., Jeanmart, H., \& Contino, F.(2021). ``\emph{Comprehensive integration of the non-energy demand within a whole-energy system: Towards a defossilisation of the chemical industry in Belgium.}'' In proceedings of  ECOS 2021 conference (Vol. 34, p. 154).
\end{quote}

\begin{quote}
Limpens, G., Coppitters, D., \textbf{Rixhon, X.}, Contino, F., \& Jeanmart, H. (2020). ``\emph{The impact of uncertainties on the Belgian energy system: application of the Polynomial Chaos Expansion to the EnergyScope model.}'' In proceedings of  ECOS 2020 conference (Vol. 33, p. 711).
\end{quote}