%!TEX root = ../thesis_main.tex
%!TEX encoding = UTF-8 Unicode

%\section{Past studies on steamd and oxygen injection in staged downdraft gasification}

Experimental studies involving simultaneous injection of air, steam and oxygen in two-stage downdraft gasifiers are rare, and often focus on the production of hydrogen rather than a high LHV and low tar content. Furthermore, none of the gasifier designs used in the reviewed studies does include a physical separation of the biomass and char beds and an intermediate oxidation zone: all of them are based on a staged air injection in a continuous porous biomass bed.

With air and saturated steam in a dual-fired downdraft gasifier, \citet{Ram2019} found that a Steam-to-Biomass mass Ratio (SBR) of around 0.2 maximized the syngas \ce{H2} content (27.2\%) and Higher Heating Value (HHV) (6.3\,MJ/Nm$^3$). \citet{Jarungthammachote2012} also investigated air-steam gasification in a two-stage gasifier, and similarly reported that SBR=0.2 yielded more \ce{H2} (up to 16.6\%) and a higher HHV (up to 5.3\,MJ/Nm$^3$) compared to the situation with air only. They simultaneously observed an increase in the tar content, explained by the temperature reduction at the location of the steam injection, which negatively affected tar thermal cracking. In these studies, the focus on HHV instead of LHV induces a bias in favour of \ce{H2}: water condensation enthalpy is usually not recovered in CHP and industrial burners, and \ce{H2} has a higher HHV but a lower LHV than \ce{CO}.

Under similar conditions with air and steam in a 50\,kW two-stage downdraft gasifier, \citet{DeSales2017} found that steam injection resulted in a higher \ce{H2} concentration (up to 20\%) but a lower syngas LHV (below 4\,MJ/Nm$^3$), due to the resulting increase in the air-fuel Equivalence Ratio (ER). They also observed that steam injection with air reduced the particle matter concentration (-50\% from 102 to 53\,mg/Nm$^3$), but increased the tar content (+69\% from 54 to 92\,mg/Nm$^3$) due to a lower temperature inside the gasifier. Using a combination of oxygen and steam, they reported a LHV up to 8.4\,MJ/Nm$^3$ and observed that the use of \ce{O2} mitigated the negative impact of steam on tar content.

\citet{Sandeep2014a} used oxygen and large amounts of superheated steam (830°C, SBR between 0.5 and 1.7) in a staged open-top gasifier, for hydrogen production. They reported a syngas LHV up to 8.9\,MJ/Nm$^3$, and observed that as the amount of steam increased, the syngas \ce{H2} content was raised up to 50\%, at the expense of a lower syngas LHV.

Only partial agreement was found on the effects of oxygen and steam on syngas LHV and tar content in staged downdraft gasifiers. The diversity of operating conditions, such as the secondary-to-primary \ce{O2} ratio and the Equivalence Ratio (ER), can explain some of the differences observed. Results are often presented in a way that is difficult to generalize to other similar installations. Overall, existing studies show that oxy-steam gasification boosts the syngas LHV, but the individual effects of oxygen and steam are not clearly distinguished as no experiment with oxygen only was found. Steam injection increases the \ce{H2} concentration, but opposite observations exist on its effect on the syngas LHV. Finally, the respective effects of oxygen and steam on tar were not studied in much depth, and seem to depend much on the temperature conditions in the gasifier.

%\citet{Ram2019} used air-steam gasification with saturated steam in a dual-fired downdraft gasifier, towards the production of \ce{H2}-enriched syngas. With values of Steam-to-Biomass mass Ratio (SBR) around 0.2, they obtained a maximum \ce{H2} content of 27.2\% and a maximum Higher Heating Value (HHV) of 6.3\,MJ/Nm$^3$. \citet{Jarungthammachote2012} also investigated air-steam gasification for hydrogen in a two-stage gasifier, and showed that compared with air-only, using SBR=0.21 yielded higher H2 (+40\%, from 11.8\% to 16.6\%) and HHV (+5.9\%, from 4.8 to 5.1\,MJ/Nm$^3$). SBR values beyond 0.21 did not lead to further improvement. They simultaneously observed a linear increase in the tar content, due to a temperature reduction in the secondary stage resulting from steam injection, which negatively affected tar thermal cracking.

%\citet{DeSales2017} reported experiments on air, air-steam and oxy-steam gasification in a 50\,kW two-stage downdraft gasifier, using saturated steam. With air and steam, the LHV and Cold Gas Efficiency (CGE) were lower than with air only, mainly due to a higher air-fuel Equivalence Ratio (ER) resulting from a lower biomass consumption rate. Both LHV and CGE were shown to decrease linearly with increasing ER. With \ce{O2} and steam, a LHV up to 8.4\,MJ/Nm$^3$ was obtained thanks to the absence of \ce{N2} dilution, an increase of +66\% compared to the highest value with air (5.0\,MJ/Nm$^3$). They also observed that steam injection with air reduced the particle matter concentration (-50\% from 102.4 to 53.4\,mg/Nm$^3$), but increased the tar concentration (+69\% from 54.25 to 91.95\,mg/Nm$^3$) due to a lower reaction temperature inside the gasifier. The latter effect was almost compensated when using oxygen instead of air, yielding a tar amount of 58--64\,mg/Nm$^3$.

%\citet{Sandeep2014a} used oxy-steam gasification in a staged gasifier for hydrogen production, using pure oxygen and large amounts of superheated steam at 830°C injected in the secondary stage (SBR between 0.48 and 1.74). They observed that as the amount of steam increased, the syngas \ce{H2} content was raised up to 50\%, at the expense of a lower syngas LHV and CGE. However, they did not compare their results with pure air gasification, and didn’t separate the individual effects of steam and oxygen.